World Heritage Sites on Wikipedia: Cultural Heritage Activism in a Context of Constrained Agency

### 2021-03-31 analysis/paper/paper.Rmd
add rrtools to dockerfile

```{r wh-wiki-cluster, fig.cap="Results of UMAP and HDBSCAN analysis to identify clusters of articles. Point colour indicates cluster membership. Inset plot shows variable loadings to show variables most influential on cluster membership", fig.width=15}
# rrtools:::quietly(source(here::here("analysis/paper/002-Wikipedia-umap-hdbscan.R")))
# knitr::include_graphics(here::here("analysis/figures/wh_country_umap_hdbscan.png"))
```

@@

The most striking spatial pattern in our results is the high proportion of anonymous edits (as the only type of edit that can be geolocated) that come from the United States on articles about CS-WHL sites, regardless of where in the world the CS-WHL site is located. These anonymous editors of articles about CS-WHL are generally not located in the same country as the site they are editing. Finding the United States at the first ranked location for anonymous edits is not surprising as it has the world's largest number of English speakers. We might similarly expect that the French language Wikipedia (2.3 m articles, 173 m edits, 3.9 m users) would have more edits originating in France or francophone on articles about CS-WHL sites in any part of the world. It is beyond the scope of this paper to analyze spatial patterns of edits in Wikipedias written in languages other than English; we focus on the English language Wikipedia because it is the largest by a considerable margin [6.2 m articles, 998 m edits, 40 m users, @wiki:listoflangs].

@@

The United States, United Kingdom, Australia and Canada, are an important group because they are collectively known as the core Anglosphere countries, sharing a common heritage as former colonies of the British empire [@Richards2019]. @vucetic2011anglosphere defines the Anglosphere as a 'distinct international, transnational, civilisational, and imperial entity within the global society'. The dominance of editors of CS-WHL articles located in the Anglosphere indicates an extension of this imperial project into the organization of world heritage information on Wikipedia. The prominence of Anglosphere editors on CS-WHL articles articles represents a digital colonisation of world heritage information on Wikipedia, where heritage is interpreted and communicated by people who are not part of the descendant communities traditionally associated with the site. Although we could only geo-locate about 20% of edits, our spatial patterns resemble those found by @Mandiberg_2020 who examined 884 m edits to the English Wikipedia and found that at the country level, the 'five largest contributors were part of what once was the British Empire, and account for nearly 75 percent of all editors'.

This is an important observation because it is these editors that control the production of knowledge about CS-WHL articles, and the ability to control information about heritage has several implications. First is that the choices of these editors define community identity, that is, the community of Wikipedia editors, and also the broader community of Internet users that read Wikipedia, by indicating which sites are important and visible, and which are not by contributing to a digital 'authorized heritage discourse' [@smith2012constrained] that prioritizes their own self-interests (cf. the poor representation of CS-WHL sites located in the Global South). Second is that traditional communities associated with CS-WHL have little control over how their own heritage is represented on Wikipedia. We see this spatial pattern on Wikipedia as part of a trend of globalization of heritage, where places that had traditionally contributed to local or regional identities are turned into sanitized playgrounds for rich tourists [@bernbeck2004political]. Wikipedia may be seen as virtual playground for Anglosphere citizens to construct a universal heritage that reflects their neo-liberal cultural values, and validates their identities, perhaps reflecting a cultural rhetoric of 'heritage populism' [@reynie2016specter]. In this way, the representation of heritage on Wikipedia recapitulates a long-standing tradition of colonial subordination of one group's heritage and identity by a more powerful and dominant group.

>>>>

```{r wh-wiki-cluster, fig.cap="Results of UMAP and HDBSCAN analysis to identify clusters of articles. Point colour indicates cluster membership. Inset plot shows variable loadings to show variables most influential on cluster membership", fig.width=15}
rrtools:::quietly(source(here::here("analysis/paper/002-Wikipedia-umap-hdbscan.R")))
# knitr::include_graphics(here::here("analysis/figures/wh_country_umap_hdbscan.png"))
```

@@

The most striking spatial pattern in our results is the high proportion of anonymous edits (as the only type of edit that can be geolocated) that come from the United States on articles about CS-WHL sites, regardless of where in the world the CS-WHL site is located. These anonymous editors of articles about CS-WHL are generally not located in the same country as the site they are editing. Finding the United States at the first ranked location for anonymous edits is not surprising as it has the world's largest number of English speakers. We might similarly expect that the French language Wikipedia (2.3 m articles, 173 m edits, 3.9 m users) would have more edits originating in France or francophone on articles about CS-WHL sites in any part of the world. We have limited the scope of our initial foray into this topic to the English language Wikipedia because it is the largest by a considerable margin [6.2 m articles, 998 m edits, 40 m users, @wiki:listoflangs]. The question of whether or not our findings generalise to non-English-language Wikipedias should be priority for future research.

@@

The United States, United Kingdom, Australia and Canada, are an important group because they are collectively known as the core Anglosphere countries, sharing a common heritage as former colonies of the British empire [@Richards2019]. @vucetic2011anglosphere defines the Anglosphere as a 'distinct international, transnational, civilisational, and imperial entity within the global society'. The dominance of editors of CS-WHL articles located in the Anglosphere indicates an extension of this imperial project into the organization of world heritage information on the English-language Wikipedia. The prominence of Anglosphere editors on CS-WHL articles articles represents a digital colonisation of world heritage information on Wikipedia, where heritage is interpreted and communicated by people who are not part of the descendant communities traditionally associated with the site. Although we could only geo-locate about 20% of edits, our spatial patterns resemble those found by @Mandiberg_2020 who examined 884 m edits to the English-language Wikipedia and found that at the country level, the 'five largest contributors were part of what once was the British Empire, and account for nearly 75 percent of all editors'.

This is an important observation because it is these editors that control the production of knowledge about CS-WHL articles, and the ability to control information about heritage has several implications. First is that the choices of these editors define community identity, that is, the community of Wikipedia editors, and also the broader community of Internet users that read Wikipedia, by indicating which sites are important and visible, and which are not by contributing to a digital 'authorized heritage discourse' [@smith2012constrained] that prioritizes their own self-interests (cf. the poor representation of CS-WHL sites located in the Global South). Second is that traditional communities associated with CS-WHL have little control over how their own heritage is represented on the English-language Wikipedia. We see this spatial pattern on Wikipedia as part of a trend of globalization of heritage, where places that had traditionally contributed to local or regional identities are turned into sanitized playgrounds for rich tourists [@bernbeck2004political]. Wikipedia may be seen as virtual playground for Anglosphere citizens to construct a universal heritage that reflects their neo-liberal cultural values, and validates their identities, perhaps reflecting a cultural rhetoric of 'heritage populism' [@reynie2016specter]. In this way, the representation of heritage on Wikipedia recapitulates a long-standing tradition of colonial subordination of one group's heritage and identity by a more powerful and dominant group.
